\normalfont\spacedallcaps{Hazardous asteroids forecast via Markov random fields}

### 2021-11-24 Manuscript/manuscript.tex
PUB

@@

The solar system is composed not only by planets and the Sun, but also by a plethora of small bodies\footnote{Small with respect to the size of planets} whose orbit can be near to the Earth \cite{burbine2016asteroids}. In particular if the orbit of these items is near less then 1.3 AU such items are called near-Earth object (NEO). Such items are of three type: comets, meteoroids and asteroids. The first one are icy bodies that moving near the sun melts and release gases giving coloured tails. Meteroids are small \footnote{With respect to asteroids} rocky/metallic items with a diameter less than 1 meter. Asteroids, on the other side, are items with a diameter larger than 1 m \cite{burbine2016asteroids}. The first two items are excluded from the present analysis.  
The near asteroids are classified, as shown in Fig. \ref{neo_orbit_types}, according to their semi-major axis, perihelion distance, and aphelion distance \cite{nasa_classification}:

\begin{itemize}
\item \textbf{Atiras}
\item \textbf{Atens}
\item \textbf{Apollos}
\item \textbf{Amors}
\end{itemize}

>>>>

The solar system is composed not only by planets and the Sun, but also by a plethora of small bodies\footnote{Small with respect to the size of planets} whose orbit can be near to the Earth \cite{burbine2016asteroids}. In particular if the orbit of these items is near less then 1.3 AU such items are called near-Earth object (NEO). Such items are of three type: comets, meteoroids and asteroids. The first one are icy bodies that moving near the sun melts and release gases giving coloured tails. Meteroids are small \footnote{With respect to asteroids} rocky/metallic items with a diameter less than 1 meter. Asteroids, on the other side, are items with a diameter larger than 1 m \cite{burbine2016asteroids}. The first two items are excluded from the present analysis.  
The near asteroids are classified, as shown in Fig. \ref{neo_orbit_types}, according to their semi-major axis (a), perihelion distance (q), and aphelion distance (Q). Here we report the Center for Near-Earth Object Studies (CNEOS) definition \cite{nasa_classification}:

\begin{itemize}
\item \textbf{Atiras} $a < 1.0$ au $Q < 0.983$ au\quad \textit{NEAs whose orbits are contained entirely with the orbit of the Earth (named after asteroid 163693 Atira)}
\item \textbf{Atens} $a < 1.0$ au $Q > 0.983$ au \quad \textit{Earth-crossing NEAs with semi-major axes smaller than Earth's (named after asteroid 2062 Aten)}
\item \textbf{Apollos} $a>1.0$ au $q<1.017$ au \quad  \textit{Earth-crossing NEAs with semi-major axes larger than Earth's (named after asteroid 1862 Apollo) }
\item \textbf{Amors} $a>1.0$ au $1.017<q<1.3$ au \textit{Earth-approaching NEAs with orbits exterior to Earth's but interior to Mars' (named after asteroid 1221 Amor)}
\end{itemize}

Beside this classification, there is the one which involve the hazardousness of an asteroid \cite{nasa_classification}:

\begin{itemize}
\item \textbf{Potentially Hazardous Asteroids}: MOID $\leq 0.05$ au $M \leq22.0$ \textit{NEAs whose Minimum Orbit Intersection Distance (MOID) with the Earth is 0.05 au or less and whose absolute magnitude (M) is 22.0 or brighter}
\end{itemize}

@@
